---
title: "Simple cepumd Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simple-cepumd-workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(cepumd)
library(gt)
```

```{r store-data-dir, echo=FALSE}
ce_data_dir <- "C:/r-projects/ce_data_dir"
```

The following is an example of using `cepumd` to calculate a 2021 annual, weighted estimate of mean expenditures on pets for all of the U.S. using CE integrated data. 

The below code assumes that CE PUMD zip files and the hierarchical grouping zip file are stored in a path the name of which is stored in a variable called `ce_data_dir` as shown below. CSV files can be acquired from the [CE PUMD Data Files page](https://www.bls.gov/cex/pumd_data.htm#csv) and hierarchical grouping files can be acquired from the [CE PUMD Documentation page](https://www.bls.gov/cex/pumd_doc.htm)
```{r show-data-dir, eval=FALSE}
ce_data_dir <- "~/Users/ce-pumd-data"
```

The first step in the workflow is to load the 2021 hierarchical grouping file for integrated expenditures using `ce_hg()`. This file is used to get the correct UCCs associated with pet expenditures.
```{r read-hg}
integ21_hg <- ce_hg(
  2021,
  integrated,
  hg_zip_path = file.path(ce_data_dir, "stubs.zip")
)
```

Next, one can use `ce_prep_data()` to correctly merge all of the Interview and Diary survey files for calculating annual estimates.
```{r prepare-data}
pet_data <- ce_prepdata(
  2021,
  integrated,
  integ21_hg,
  uccs = ce_uccs(integ21_hg, expenditure = "Pets", ucc_group = "PETS"),
  dia_zp = file.path(ce_data_dir, "diary21.zip"),
  int_zp = c(
    file.path(ce_data_dir, "intrvw20.zip"),
    file.path(ce_data_dir, "intrvw21.zip")
  )
)

gt(head(pet_data))
```

The last step is to run `ce_mean()` to get annual, weighted pet expenditure estimates.
```{r calculate-estimate}
ce_mean(pet_data) |>
  gt()
```

